The transiting dust clumps in the evolved disk of the Sun-like UXor RZ~Psc

### 2016-12-09 ms.tex
preparing for resubmission

A promising way to probe these inner regions is by observing temporal variability
(e.g. \cite{2014Sci...345.1032M}). Optical and IR stellar variation has been studied for
decades (e.g. \cite{1945ApJ...102..168J,1994AJ....108.1906H}), and has recently been
reinvigorated by large scale efforts
(e.g. \cite{2011ApJ...733...50M,2014AJ....147...82C}) and as a side-effect of large-scale
surveys for transiting planets \cite{2012AJ....143...72M,2013AJ....146..112R}. Of many
different classes of variables, the ones of most interest and relevance here are the
``UXors'', named for the prototypical system UX~Orionis \cite{1994AJ....108.1906H}. These
are usually Herbig Ae and late-type Herbig Be stars \cite{1984A&AS...55..109F}, and
typically show several magnitudes of extinction that is generally attributed to variable
obscuration by circumstellar dust
\cite{1994AJ....108.1906H,1999AJ....118.1043H,2000A&A...364..633N}.\footnote{Not all
  stars occulted by dust are UXors. Two main classes are: those occulted by circumstellar
  material beyond $\gtrsim$10au, such as AA~Tau and V409~Tau which show
  $\gtrsim$year-long dimming events \cite{2013A&A...557A..77B,2015AJ....150...32R}, and
  systems such as $\epsilon$~Aur, J1407, EE~Cep, and OGLE-LMC-ECL-11893 where the
  occultations are attributed to circumsecondary disks
  (e.g. \cite{1999MNRAS.303..521M,2012AJ....143...72M,2014ApJ...788...41D,2015ApJS..220...14K}).}
Two key arguments that favour circumstellar dust as the cause are i) ``blueing'', where
the star is reddened when moderately dimmed but returns to the stellar colour for very
deep ($\gtrsim$1mag) events -- the remaining flux is interpreted as light scattered off
circumstellar dust \cite{1988SvAL...14...27G}, and ii) increased polarisation fraction
during dimming events, caused by a greater fraction of the flux being contributed by
dust-scattered light (e.g. \cite{2001ARep...45...51R}, which also shows that the
surrounding dust does not reside in a spherical shell). UXors therefore reveal
information on the degree of non-axisymmetry, the ``clumpiness'', of dust orbiting a star
on spatial scales similar to the star itself. The observations can span multiple orbits
to test for repeated dimming events (e.g. \cite{1999A&A...349..619B}), and by using
different bandpasses and polarisation can estimate dust grain sizes
(e.g. \cite{1988SvAL...14...27G,1994ASPC...62...63G}).

@@

Given an interpretation related to individual planetesimal disruptions, rather than
hydrodynamics, it is perhaps surprising that to date the dimming events are not seen to
be periodic \cite{1999AstL...25..243R,2013A&A...553L...1D}. The only cyclical variation
seen in light curves for RZ~Psc is a 12.4 year variation with an amplitude of 0.5 mag,
which is attributed to either a magnetic cycle, or precession of an otherwise unseen
outer disk due to perturbations from an unseen companion \cite{2013A&A...553L...1D}.

@@

To study the temporal variability of RZ~Psc we use two seasons of public data from the
Wide-Angle Search for Planets (WASP \cite{2006PASP..118.1407P,2010A&A...520L..10B}), and
nine seasons of data from the Kilodegree Extremely Little Telescope North (KELT-North
\cite{2007PASP..119..923P}). We also collected, but ultimately did not use, photometric
observations of RZ~Psc from a wide variety of other sources
(\cite{1994AJ....108.1906H,1973IBVS..783....1K,1980PZ.....21..310K,1985PZ.....22..181Z,1991Afz....34..333K,1997AcA....47..467P,2014Ap.....57..491P},
the Catalina Sky Survey, the American Association of Variable Star Observers, the All-Sky
Automated Survey). Aside from the Harvard plate photometry published by G\"urtler et
al. \cite{1999A&AS..140..293G} we have not sought unpublished photometry so the light
curve remains incomplete.\footnote{Normalised photometry is available at
  \href{https://github.com/drgmk/rzpsc}{https://github.com/drgmk/rzpsc}}

@@

Motivated by this variation, we obtained VLT/VISIR observations of RZ~Psc; an N-band
(10$\mu$m) spectrum on 2016 August 16 and Q-band (18$\mu$m) photometry on 2016 July 27
(programme 097.C-0217). These data, and the related calibration observations, were
reduced using the standard ESO \texttt{esorex} pipeline. The wall-clock integration time
for the spectrum was 50 minutes at an airmass of 1.66, and was calibrated using an
observation of HD~189831 taken immediately afterwards at an airmass of 1.63. The spectra
in individual chop/nod cycles are consistent, suggesting that the absolute calibration is
reliable. The spectrum was trimmed to mask highly uncertain regions shortward of 8$\mu$m,
and longward of 13 $\mu$m, and near the telluric absorption at 9.5$\mu$m. The Q-band
photometry took 45 minutes at an airmass of 1.65-1.7, and was calibrated against an
observation of HD~2436 taken immediately afterwards at an airmass of 1.5. Photometry of
RZ~Psc and HD~2436 was done using a 0.9'' radius aperture and a sky annulus from
1-2''. In addition to the conversion from adu/s to Jansky using HD~2436, we applied an
additional upward correction to account for the slightly lower airmass for the calibrator
($\exp(0.3 [1.675-1.5] \approx 1.05$, where an extinction of 0.3 per unit airmass was
assumed). Uncertainties were estimated as the standard deviation of the flux density in
50 apertures around RZ~Psc. The final calibrated flux is $86 \pm 10$ mJy.

@@

To further explore this variability, we use the WISE ``epoch'' photometry, for which the
telescope scanning strategy results in clusters of measurements that are spaced 6 months
apart. These data appear at approximately days 70 and 250 in the relevant years in Figure
\ref{fig:waspkelt}, but do not coincide with any optical dimming events. Photometry is
not available in all four channels since launch in early 2010, due to exhaustion of the
coolant after 7.7 months and a 2.5-year hiatus from mid 2011-2014 (see
\cite{2010AJ....140.1868W,2014ApJ...792...30M}). These data are shown in Figure
\ref{fig:wisevar}, where the left panel shows the 3.4$\mu$m magnitude as a function of
time, and that there is significant variation on 6-month timescales, and an even greater
variation overall. Inspection of the individual clusters, which are on hour to day
timescales, shows no significant variation with time. The dashed lines have the slope
expected for disk brightness variation that is independent of wavelength; the slopes are
not exactly 1 because the total flux is not dominated by the disk near 3-5$\mu$m, and
hence the slopes are slightly flatter. Comparing the observed and expected correlations,
we conclude that the data do not show significant evidence for changes in the spectral
shape (i.e. changing temperature or composition). However, the ratios including 12 and
22$\mu$m observations are most sensitive to these change, but only include the first two
sets of observations where the brightness changes were relatively small.

@@

The coverage in an individual season is 100-150 days. Thus, if the occulting material
resides in an asteroid belt closer than $\sim$0.5au, periodicity in the dimming events
may be visible in a single season's data. Longer orbital periods may be visible across
seasons, though the six-month gap between seasons makes unambiguously linking events
harder. Non-detection of periodicity would imply that the data are not sufficient, or
that strict periodicity does not exist. An intermediate possibility is that occultations
happen with a range of periodicities, perhaps reflecting their origin in a radially broad
region, and that discerning this scenario from randomly occurring occultations is not
possible given the data.

@@

We first used autocorrelation to search for periodicity, rather than methods related to
Fourier transforms (e.g. periodograms). The motivation being that an individual transit
event may be followed by another some number of days later, and perhaps repeat a few
times, but other similarly (but not exactly) separated events may happen years later or
earlier with a phase that is totally different. We therefore used the discrete
autocorrelation function (DACF) proposed by \cite{1988ApJ...333..646E}, though do not
include uncertainties on individual measurements. For a time series with measurements
$a_i$ at times $t_i$ the DACF first computes the mean $\bar{a}$ from the light curve,
here we used the sigma-clipped mean to remove the dimming events. Then for each pair of
points $a_i$, $a_j$ (with $i\ne j$) computes
$U_{ij}=((a_i-\bar{a})(a_j-\bar{a}) / \sigma_a$, with each $U_{ij}$ associated with a
time lag $\Delta t_{ij}=t_j - t_i$. A series of time lags centered at times $t_{\rm lag}$
with width $\Delta t_{\rm lag}$ are then used as bins, and the average in each bin is the
DACF. The DACF is not computed for lag bins with no data. The units of the DACF are
standard deviations of the light curve $\sigma_a$ (again calculated using sigma
clipping).

@@

The solid lines in Figure \ref{fig:hist} shows this power in histogram form (the same in
each panel), generated from 60 events that were identified in the light curve. The y-axis
should be interpreted as the number of events that are consistent with that period on the
x-axis. As before, there is evidence for a peak, now slightly shifted to near 65 days. In
contrast to the DACF analysis, this power does not all arise from a single year. Most is
contributed by 2004 and 2009, but exclusion of events from these years results in a
similar (but noisier) histogram.

To test what could have been detected, and the quantify the variation in power expected,
we created synthetic light curves with the same temporal sampling as the WASP and
KELT-North data. To do this we set the flux to 1, and randomly injected a number of
dimming events with a flux of 0.1. These events are therefore relatively easy for the
algorithm to detect, but suffer the same sampling issues. We tried individual randomly
occurring events, and periodic events that repeat a fixed number of times. All events
have durations randomly distributed between 1 and 4 days, similar to the observed
events. For random events the remaining parameter in this model is simply the number of
events -- this is the total number over the 11 year period covered by the WASP and
KELT-North data, including when measurements were not being taken. For repeating events
the range of periods and the number of repeats are additional parameters. To estimate the
level of variation in the power spectrum we repeated the process of injecting synthetic
events 500 times, and in each bin estimate the mean and standard deviation of the power.

@@

Using the simple formalism described above, we can use gradients derived directly from
the light curves to estimate the radial location of the occulting clumps. The gradients
are estimated by least squares fitting straight lines to each night's observations, for
which only nights with six or more measurements are used. This procedure is possible
because in nearly all cases an individual night's observations only cover ingress or
egress, not both. These gradients are plotted against the minimum nightly flux
$f_{\rm min}$ in Figure \ref{fig:grad}. We plot gradients whose uncertainty is less than
4$\times$ their value as open circles, with symbol sizes proportional to the inverse of
this uncertainty. All other gradients are plotted as small dots, as a check that the
gradients for unocculted fluxes are near zero. The horizontal scatter of these gradients
near $f_{\rm min}=0$ provides a further estimate of the uncertainty of individual
gradients. The solid lines show the gradients expected for circular orbits at a range of
semi-major axes.

@@

In Figure \ref{fig:sma} we have converted the gradients into semi-major axes using
equation (\ref{eq:acl}). The dashed line shows the stellar radius, estimated to be
approximately Solar based on an age of 25Myr \cite{2013AstL...39..776P} and an effective
temperature of 5350K \cite{2014A&A...563A.139P} using Siess \cite{2000A&A...358..593S}
isochrones. This Figure also includes solid lines of constant gradient, computed using
equation (\ref{eq:acl}) and $f_{\rm min} = 1-\tau$.

The blue line shows the semi-major axis implied by the scatter in gradients near
$f_{\rm min}=0$ in Figure \ref{fig:grad}. Above this line the gradients can be considered
consistent with zero, and the inferred semi-major axes largely meaningless. A related
point is that because the points near $f_{\rm min}=0$ have very small $\tau$, the derived
velocities and semi-major axes can be rather extreme and should be disregarded. An
additional issue is highlighted by the two large circles near $f_{\rm min}=0.5$ and
between $a_{\rm circ}=30$ to 100, which are the first two nights of 2006 observations
shown in Figure \ref{fig:waspkeltzoom}. The first is at the time of minimum flux, and the
other also appears to be at a turning point, so the assumption behind equation
(\ref{eq:acl}) that $u=0$ (i.e. the cloud edge is passing the stellar disk center) is
clearly incorrect. These points should therefore be associated not with clumps at 30 to
100au, with clumps that lie somewhere interior.

The green line shows the semi-major axis implied by a gradient of 1/day. As the lowest
flux measured on a given night is not necessarily the minimum flux for that dimming
event, the inferred semi-major axis could be greater; in this case points move downward
parallel to this and similar lines. Thus, the main constraint offered by Figure
\ref{fig:sma} is that the clumps orbit with semi-major axes less than 10au.

Figure \ref{fig:sma} suggests that the clumps orbit with projected velocities that are
consistent with circular orbits at about 1au. The conclusion from the IR excess, and
possibly the periodicity analysis, is that they may lie closer, near
0.3-0.7au. Reconciling these differences requires that the clumps are on eccentric
obits. For a semi-major axis of 0.3au the minimum clump eccentricity (i.e. all clumps
transit at apocenter) is about 0.6. Such an orientation is of course highly unlikely, so
either the semi-major axis is larger, or the true eccentricities need to be
higher. Perhaps coincidentally, an eccentricity of 0.6 yields a pericenter velocity of
120km s$^{-1}$, similar to the velocity shifts in Sodium absorption lines seen towards
RZ~Psc, which are discussed further in section \ref{s:disk}\ref{ss:orig} (noting however
that the pericenter velocity is tangential, and the absorption lines show projected
radial velocity \cite{2013Ap.....56..453P}).

@@

The primary uncertainty lies with the disk evolutionary state. If a significant gas
reservoir remains, a turbulent inner rim scenario similar to that proposed for UXors
could produce a similar light curve. Both the lack of accretion, and a likely age beyond
which gas-rich disks are typically seen, may argue against this scenario, though the disk
may be in transition to the debris phase and retain some primordial gas. We revisit the
disk status from the perspective of the flux distribution in section
\ref{s:disk}\ref{ss:evol}.

@@

We now address several open questions, each taking a slightly wider view. Primary among
these is whether the dimming events and the IR excess are caused by the same dust, as
suggested by \cite{2013A&A...553L...1D}. Two further aspects are then the implications
for the origin of the clumps and the evolutionary status of the disk in which they
reside. We finish by considering the proposed scenarios for dippers and UXors, and why
RZ~Psc appears to be a rare object that lies between these classes.

@@

To explore possible geometries we use a simple relation between fractional luminosity
$L_{\rm disk}/L_\star$, (uniform) geometric optical depth $\tau$, and the disk opening
angle $\theta$ \cite{2014MNRAS.438.3299K}
\begin{equation}
  L_{\rm disk}/L_\star = \tau sin(\theta/2) \, ,
\end{equation}
which says that the fractional luminosity is the optical depth of the dust multiplied by
the fraction of the sky covered as seen from the star. The dust belt must therefore have
an opening angle of at least 8$^\circ$ to capture 7\% of the starlight. However, for this
minimal estimate the dust is optically thick, yet RZ~Psc is not seen to be reddened. If
we instead require $\tau \sim \bar{E}$ then as stated in the previous paragraph the dust
distribution must instead be near isotropic. Given the ubiquity of disk like structures
around young stars, such a spherical distribution seems physically unlikely. In addition,
the increased polarisation during deep dimming events argues against a spherical
distribution.

Thus, the picture of RZ~Psc as a star seen through a \emph{disk}, where the clumps
account for all of the dust and sample some representative part of an asteroid belt
(e.g. the midplane), is untenable because that belt would cause much more reddening than
is observed. These characteristics distinguish RZ~Psc from heavily reddened objects,
where the dimming events could be sampling a more representative section of the disk
(e.g. \cite{2015MNRAS.451...26S}). This issue can be avoided by invoking a spherical
distribution of material, but the problem then shifts to whether such a distribution is
physically plausible.

A more likely alternative, which we favour, is that most of the dust does not lie on
orbits that pass in front of the star, and the occultations are caused by a small
fraction of objects that have greater orbital inclinations than average. In this case the
component that causes most of the IR excess may or may not be clumpy, and could be
radially optically thick (i.e. the opening angle could be as small as 8$^\circ$). This
picture unfortunately loses any strong connection between the occulting clumps and the IR
excess, essentially adding a free parameter that is the fraction of material that is
``kicked'' or resides above the disk, but seems to be the simplest and most probable
scenario. Dullemond et al (2003) used essentially the same argument for UXors
\cite{2003ApJ...594L..47D}, so our picture is therefore inevitably similar to dust
occultation models proposed for UXors
(e.g. \cite{1997ApJ...491..885N,2000A&A...364..633N,2003ApJ...594L..47D}).  As the disk
is probably radially optically thick with a scale height similar to gas-rich
protoplanetary disks, it could be that the scenario for RZ~Psc is in fact the same as
proposed for UXors. In this case, the IR excess would originate from the inner edge of a
more extended disk, which is not detected at longer wavelengths for reasons discussed
below.

We therefore conclude that while there is almost certainly some connection between the
dimming events and the IR excess, it is at best indirect; we are (unfortunately) not
viewing RZ~Psc through a representative part of an asteroid belt analogue. As with other
UXors, a clear prediction is that the disk is not seen edge-on, but at an intermediate
inclination.

@@

A further difference between RZ~Psc and UXors is the origin of the dust location. For
Herbig Ae/Be stars (and by extension, UXors) the dust inner radius is set by
sublimation. However, the dust around RZ~Psc is roughly 500K, so much cooler than the
$\sim$1500K sublimation temperature. That is, if the origin of RZ~Psc's variability is
interpreted as similar to other UXors and originates in a gas-rich disk, it must host a
transition disk rather than a ``full'' primordial disk. Therefore, while these
comparisons show RZ~Psc to be unusual compared to typical UXors, they do not argue
strongly for or against a specific scenario. We discuss the evolutionary state of the
disk further below.

>>>>

A promising way to probe these inner regions is by observing temporal variability
(e.g. \cite{2014Sci...345.1032M}). Optical and IR stellar variation has been studied for
decades (e.g. \cite{1945ApJ...102..168J,1994AJ....108.1906H}), and has recently been
reinvigorated by large scale efforts
(e.g. \cite{2011ApJ...733...50M,2014AJ....147...82C}) and as a side-effect of large-scale
surveys for transiting planets \cite{2012AJ....143...72M,2013AJ....146..112R}. Of many
different classes of variables, the ones of most interest and relevance here are the
``UXors'', named for the prototypical system UX~Orionis \cite{1994AJ....108.1906H}. These
are usually Herbig Ae and late-type Herbig Be stars \cite{1984A&AS...55..109F}, and
typically show several magnitudes of extinction that is generally attributed to variable
obscuration by circumstellar dust
\cite{1994AJ....108.1906H,1999AJ....118.1043H,2000A&A...364..633N}.\footnote{Not all
  stars occulted by dust are UXors. Two other classes are: those occulted by
  circumstellar material beyond $\gtrsim$10au, such as AA~Tau and V409~Tau which show
  $\gtrsim$year-long dimming events \cite{2013A&A...557A..77B,2015AJ....150...32R}, and
  systems such as $\epsilon$~Aur, J1407, EE~Cep, and OGLE-LMC-ECL-11893 where the
  occultations are attributed to circumsecondary disks
  (e.g. \cite{1999MNRAS.303..521M,2012AJ....143...72M,2014ApJ...788...41D,2015ApJS..220...14K}).}
Three related arguments that favour circumstellar dust as the cause are i) a maximum
depth of dimming events of roughly 3 magnitudes, suggesting that a few percent of the
visible flux is not directly from the star, but scattered off a disk that surrounds the
star and remains visible even when the star itself is completely occulted, ii)
``blueing'', where the star is reddened for small ($\lesssim$1mag) levels of dimming but
returns to the stellar colour (i.e. becomes ``bluer'') for the very deep ($\gtrsim$1mag)
events where the star is mostly occulted -- the reddening indicates dimming by
circumstellar dust, and a stellar colour is typical of light scattered off circumstellar
dust \cite{1988SvAL...14...27G}, and iii) increased polarisation fraction during dimming
events, caused by a greater fraction of the flux being contributed by dust-scattered
light (e.g. \cite{2001ARep...45...51R}, which also shows that the surrounding dust does
not reside in a spherical shell). UXors therefore reveal information on the degree of
non-axisymmetry, the ``clumpiness'', of dust orbiting a star on spatial scales similar to
the star itself. The observations can span multiple orbits to test for repeated dimming
events (e.g. \cite{1999A&A...349..619B}), and by using different bandpasses and
polarisation can estimate dust grain sizes
(e.g. \cite{1988SvAL...14...27G,1994ASPC...62...63G}).

@@

Given the proposed interpretation related to individual planetesimal disruptions, rather
than hydrodynamics, it is perhaps surprising that to date the dimming events are not seen
to be periodic \cite{1999AstL...25..243R,2013A&A...553L...1D}. The only cyclical
variation seen in light curves for RZ~Psc is a 12.4 year variation with an amplitude of
0.5 mag, which is attributed to either a magnetic cycle, or precession of an otherwise
unseen outer disk due to perturbations from an unseen companion
\cite{2013A&A...553L...1D}.

@@

To study the temporal variability of RZ~Psc we use two seasons of public data from the
Wide-Angle Search for Planets (WASP \cite{2006PASP..118.1407P,2010A&A...520L..10B}), and
nine seasons of data from the Kilodegree Extremely Little Telescope North (KELT-North
\cite{2007PASP..119..923P}). We also collected, but ultimately did not use, photometric
observations of RZ~Psc from a wide variety of other sources
(\cite{1994AJ....108.1906H,1973IBVS..783....1K,1980PZ.....21..310K,1985PZ.....22..181Z,1991Afz....34..333K,1997AcA....47..467P,2014Ap.....57..491P},
the Catalina Sky Survey, the American Association of Variable Star Observers, the All-Sky
Automated Survey). Aside from the Harvard plate photometry published by G\"urtler et
al. \cite{1999A&AS..140..293G} we have not sought unpublished photometry so the light
curve remains incomplete.\footnote{All normalised photometry is available at
  \href{https://github.com/drgmk/rzpsc}{https://github.com/drgmk/rzpsc}}

@@

Motivated by this variation, we obtained VLT/VISIR observations of RZ~Psc; an N-band
(10$\mu$m) spectrum on 2016 August 16 and Q-band (18$\mu$m) photometry on 2016 July 27
(programme 097.C-0217). These data, and the related calibration observations, were
reduced using the standard ESO \texttt{esorex} pipeline. The wall-clock integration time
for the spectrum was 50 minutes at an airmass of 1.66, and was calibrated using an
observation of HD~189831 taken immediately afterwards at an airmass of 1.63. The spectra
in individual chop/nod cycles are consistent so the shape of the spectrum is
reliable. The absolute calibration is uncertain at $\sim$10\% levels
\cite{2007A&A...476..279G}, which is sufficiently precise for our purposes here. The
spectrum was trimmed to mask highly uncertain regions shortward of 8$\mu$m, and longward
of 13 $\mu$m, and near the telluric absorption at 9.5$\mu$m. The Q-band photometry took
45 minutes at an airmass of 1.65-1.7, and was calibrated against an observation of
HD~2436 taken immediately afterwards at an airmass of 1.5. Photometry of RZ~Psc and
HD~2436 was done using a 0.9'' radius aperture and a sky annulus from 1-2''. In addition
to the conversion from adu/s to Jansky using HD~2436, we applied an additional upward
correction to account for the slightly lower airmass for the calibrator
($\exp(0.3 [1.675-1.5] \approx 1.05$, where an extinction of 0.3 per unit airmass was
used, derived from archival calibration data using a method similar to that of Verhoeff
et al. \cite{2012A&A...538A.101V}). Uncertainties were estimated as the standard
deviation of the flux density in 50 apertures around RZ~Psc. The final calibrated flux is
$86 \pm 10$ mJy.

@@

To further explore this variability, we use the WISE ``epoch'' photometry, for which the
telescope scanning strategy results in clusters of measurements that are spaced 6 months
apart. These data appear at approximately days 70 and 250 in the relevant years in Figure
\ref{fig:waspkelt}, but do not coincide with any optical dimming events. Photometry is
not available in all four channels since launch in early 2010, due to exhaustion of the
coolant after 7.7 months and a 2.5-year hiatus from mid 2011-2014 (see
\cite{2010AJ....140.1868W,2014ApJ...792...30M}). These data are shown in Figure
\ref{fig:wisevar}, where the left panel shows the 3.4$\mu$m magnitude as a function of
time, and that there is significant variation on 6-month timescales, and an even greater
variation overall. Inspection of the individual clusters, which are on hour to day
timescales, shows no significant variation with time. The dashed lines have the slope
expected for disk brightness variation that is independent of wavelength; the slopes are
not exactly 1 because the total flux is not dominated by the disk near 3-5$\mu$m, and
hence the slopes are slightly flatter. Comparing the observed and expected correlations,
we conclude that the data do not show significant evidence for changes in the spectral
shape (i.e. changing temperature or composition). However, the ratios including 12 and
22$\mu$m observations are most sensitive to these changes, but only include the first two
sets of observations where the brightness changes were relatively small.

@@

The temporal coverage of the observations in an individual season is 100-150 days. Thus,
if the occulting material resides in an asteroid belt closer than $\sim$0.5au,
periodicity in the dimming events may be visible in a single season's data. Longer
orbital periods may be visible across seasons, though the six-month gap between seasons
makes unambiguously linking events harder. Non-detection of periodicity would imply that
the data are not sufficient, or that strict periodicity does not exist. An intermediate
possibility is that occultations happen with a range of periodicities, perhaps reflecting
their origin in a radially broad region, and that discerning this scenario from randomly
occurring occultations is not possible given the data.

@@

We first used autocorrelation to search for periodicity, rather than methods related to
Fourier transforms (e.g. periodograms). The motivation being that an individual transit
event may be followed by another some number of days later, and perhaps repeat a few
times, but other similarly (but not exactly) separated events may happen years later or
earlier with a phase that is totally different. We therefore used the discrete
autocorrelation function (DACF) proposed by Edelson \& Krolik \cite{1988ApJ...333..646E},
though do not include uncertainties on individual measurements. For a time series with
measurements $a_i$ at times $t_i$ the DACF first computes the mean $\bar{a}$ from the
light curve, here we used the sigma-clipped mean to remove the dimming events. Then for
each pair of points $a_i$, $a_j$ (with $i\ne j$) computes
$U_{ij}=((a_i-\bar{a})(a_j-\bar{a}) / \sigma_a$, with each $U_{ij}$ associated with a
time lag $\Delta t_{ij}=t_j - t_i$. A series of time lags centered at times $t_{\rm lag}$
with width $\Delta t_{\rm lag}$ are then used as bins, and the average in each bin is the
DACF. The DACF is not computed for lag bins with no data. The units of the DACF are
standard deviations of the light curve $\sigma_a$ (again calculated using sigma
clipping).

@@

The solid lines in Figure \ref{fig:hist} show this power in histogram form (the same in
each panel), generated from 60 events that were identified in the light curve. The y-axis
should be interpreted as the number of events that are consistent with that period on the
x-axis. As before, there is evidence for a peak, now slightly shifted to near 65 days. In
contrast to the DACF analysis, this power does not all arise from a single year. Most is
contributed by 2004 and 2009, but exclusion of events from these years results in a
similar (but noisier) histogram.

To test what could have been detected, and quantify the variation in power expected,
we created synthetic light curves with the same temporal sampling as the WASP and
KELT-North data. To do this we set the flux to 1, and randomly injected a number of
dimming events with a flux of 0.1. These events are therefore relatively easy for the
algorithm to detect, but suffer the same sampling issues. We tried individual randomly
occurring events, and periodic events that repeat a fixed number of times. All events
have durations randomly distributed between 1 and 4 days, similar to the observed
events. For random events the remaining parameter in this model is simply the number of
events -- this is the total number over the 11 year period covered by the WASP and
KELT-North data, including when measurements were not being taken. For repeating events
the range of periods and the number of repeats are additional parameters. To estimate the
level of variation in the power spectrum we repeated the process of injecting synthetic
events 500 times, and in each bin estimate the mean and standard deviation of the power.

@@

Using the simple formalism described above, we can use gradients derived directly from
the light curves to estimate the radial location of the occulting clumps. The gradients
are estimated by least squares fitting straight lines to each night's observations, for
which only nights with six or more measurements are used. This procedure is possible
because in nearly all cases an individual night's observations only cover ingress or
egress, not both. These gradients are plotted against the minimum nightly flux
$f_{\rm min}$ in Figure \ref{fig:grad}. We plot gradients whose uncertainty is less than
4$\times$ their value as open circles, with symbol sizes proportional to the inverse of
this uncertainty. All other gradients are plotted as small dots, as a check that the
gradients for unocculted fluxes are near zero. The horizontal scatter of these gradients
near $f_{\rm min}=1$ provides a further estimate of the uncertainty of individual
gradients. The solid lines show the gradients expected for circular orbits at a range of
semi-major axes.

@@

In Figure \ref{fig:sma} we have converted the gradients into semi-major axes using
equation (\ref{eq:acl}). The dashed line shows the stellar radius, estimated to be
approximately Solar based on an age of 25Myr \cite{2013AstL...39..776P} and an effective
temperature of 5350K \cite{2014A&A...563A.139P} using Siess et
al. \cite{2000A&A...358..593S} isochrones. This Figure also includes solid lines of
constant gradient, computed using equation (\ref{eq:acl}) and $f_{\rm min} = 1-\tau$.

The blue line shows the semi-major axis implied by the scatter in gradients near
$f_{\rm min}=1$ in Figure \ref{fig:grad}. Above this line the gradients can be considered
consistent with zero, and the inferred semi-major axes largely meaningless. A related
point is that because the points near $f_{\rm min}=1$ have very small $\tau$, the derived
velocities and semi-major axes can be rather extreme and should be disregarded. An
additional issue is highlighted by the two large circles near $f_{\rm min}=0.5$ and
between $a_{\rm circ}=30$ to 100 au, which are the first two nights of 2006 observations
shown in Figure \ref{fig:waspkeltzoom}. The first is at the time of minimum flux, and the
other also appears to be at a turning point, so the assumption behind equation
(\ref{eq:acl}) that $u=0$ (i.e. the cloud edge is passing the stellar disk center) is
clearly incorrect. These points should therefore be associated not with clumps at 30 to
100au, but with clumps that lie somewhere interior.

The green line shows the semi-major axis implied by a gradient of 1/day. As the lowest
flux measured on a given night is not necessarily the minimum flux for that dimming
event, the inferred semi-major axis could be greater; in this case points move downward
parallel to this and similar lines. Figure \ref{fig:sma} suggests that the clumps orbit
with projected velocities that are consistent with circular orbits at or beyond about
1au. The conclusion from the IR excess, and possibly the periodicity analysis, is that
they may lie closer, near 0.3-0.7au. Reconciling these differences requires that the
clumps are on eccentric obits. For a semi-major axis of 0.3au the minimum clump
eccentricity (i.e. all clumps transit at apocenter) is about 0.6. Such an orientation is
of course highly unlikely, so either the semi-major axis is larger, or the true
eccentricities need to be higher. Perhaps coincidentally, an eccentricity of 0.6 yields a
pericenter velocity of 120km s$^{-1}$, similar to the velocity shifts in Sodium
absorption lines seen towards RZ~Psc, which are discussed further in section
\ref{s:disk}\ref{ss:orig} (noting however that the pericenter velocity is tangential, and
the absorption lines show projected radial velocity \cite{2013Ap.....56..453P}).

@@

The primary uncertainty lies with the disk evolutionary state. If a significant gas
reservoir remains, a turbulent inner rim scenario similar to that proposed for UXors
might produce a similar light curve. Both the lack of accretion, and a likely age beyond
which gas-rich disks are typically seen, may argue against this scenario, though the disk
may be in transition to the debris phase and retain some primordial gas. We revisit the
disk status from the perspective of the flux distribution in section
\ref{s:disk}\ref{ss:evol}.

@@

We now address several open questions, each taking a slightly wider view. Primary among
these is whether the dimming events and the IR excess are caused by the same dust, as
suggested by de Wit et al. \cite{2013A&A...553L...1D}. Two further aspects are then the
implications for the origin of the clumps and the evolutionary status of the disk in
which they reside. We finish by considering the proposed scenarios for dippers and UXors,
and why RZ~Psc appears to be a rare object that lies between these classes.

@@

To explore possible geometries we use a simple relation between fractional luminosity
$L_{\rm disk}/L_\star$, (uniform) geometric optical depth $\tau$, and the disk opening
angle $\theta$ \cite{2014MNRAS.438.3299K}
\begin{equation}
  L_{\rm disk}/L_\star = \tau sin(\theta/2) \, ,
\end{equation}
which says that the fractional luminosity is the optical depth of the dust multiplied by
the fraction of the sky covered as seen from the star. The dust belt must therefore have
an opening angle of at least 8$^\circ$ to capture 7\% of the starlight. However, for this
minimal estimate the dust is optically thick, yet RZ~Psc is not seen to be reddened. If
we instead require $\tau \sim \bar{E}$ then as stated in the previous paragraph the dust
distribution must instead be near isotropic. Given the ubiquity of disk-like structures
around young stars, such a spherical distribution seems physically unlikely. In addition,
the increased polarisation during deep dimming events argues against a spherical
distribution.

Thus, the picture of RZ~Psc as a star seen \emph{through} a disk, where the clumps
account for all of the dust and sample some representative part of an asteroid belt
(e.g. the midplane), is untenable because that belt would cause much more reddening than
is observed. These characteristics distinguish RZ~Psc from heavily reddened objects,
where the dimming events could be sampling a more representative section of the disk
(e.g. \cite{2015MNRAS.451...26S}). This issue can be avoided by invoking a spherical
distribution of material, but the problem then shifts to whether such a distribution is
physically plausible.

A more likely alternative, which we favour, is that most of the dust does not lie on
orbits that pass in front of the star, and the occultations are caused by a small
fraction of objects that have higher vertical locations (or greater orbital inclinations)
than average. In this case the component that causes most of the IR excess may or may not
be clumpy, and could be radially optically thick (i.e. the opening angle could be as
small as 8$^\circ$). This picture unfortunately loses any strong connection between the
occulting clumps and the IR excess, essentially adding a free parameter that is the
fraction of material that is ``kicked'' or resides above the disk, but seems to be the
simplest and most probable scenario. Dullemond et al. \cite{2003ApJ...594L..47D} used
essentially the same argument for UXors, so our picture is therefore inevitably similar
to dust occultation models proposed for UXors
(e.g. \cite{1997ApJ...491..885N,2000A&A...364..633N,2003ApJ...594L..47D}).  As the disk
is probably radially optically thick with a scale height similar to gas-rich
protoplanetary disks, it could be that the scenario for RZ~Psc is in fact the same as
proposed for UXors. In this case, the IR excess would originate from the inner edge of a
more extended disk, which is not detected at longer wavelengths for reasons discussed
below.

We therefore conclude that while there is almost certainly some connection between the
dimming events and the IR excess, it is at best indirect; we are not viewing RZ~Psc
through a representative part of an asteroid belt analogue. As with other UXors, a clear
prediction is that the disk is not seen edge-on, but at an intermediate inclination.

@@

A further difference between RZ~Psc and UXors is the origin of the dust location. For
Herbig Ae/Be stars (and by extension, UXors) the dust inner radius is set by
sublimation. However, the dust around RZ~Psc is roughly 500K, so much cooler than the
$\sim$1500K sublimation temperature. That is, if the origin of RZ~Psc's variability is
interpreted as similar to other UXors and originates in a gas-rich disk, it must host a
transition disk rather than a ``full'' primordial disk. Therefore, while these
comparisons show RZ~Psc to be unusual compared to typical UXors, they do not argue
strongly for or against a specific scenario.
